\textsf{Are large language models superhuman chemists?}

### 2024-10-24 src/tex/ms.tex
feat: incorporate feedback from Macjonathan and Micha (#66)

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that LLMs' user base is wider than that of chemistry and materials science experts who can critically reflect on every output these models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts \autocite{wu2018moleculenet}.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards \ref{sec:leaderboard}, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions). In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations). We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human participants answered all the questions in this subset.

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts. Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo. This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
This indicates that there is still room for improvement in chemical \gls{llm}s by training them on more specialized data sources or by integrating them with specialized databases.

In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}). This is in line with observations in other domains but also indicates that chemical \gls{llm}s could, to some extent, be further improved by scaling them up.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:barplot_all_correct_all_questions}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).

 These findings also shine an interesting light on the value of textbook-inspired questions.
 A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students.
 On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
 For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
 While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.

@@

% Estimate of difficulty
\paragraph{Confidence estimates} One might wonder whether the models can estimate if they can answer a question correctly.
If they could do so, incorrect answers would be less problematic as one could detect when an answer is incorrect.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

These questions assess the ability to establish a \enquote{preference}, such as favoring a specific molecule. Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that lasts several years within a chemist's career.
Our data corpus is adapted from the published dataset by \textcite{Choung_2023}, which consists of more than 5000 question-answer pairs about chemical intuition. To build the dataset, they presented 35 medicinal chemists with two different molecules, asking them what molecule they would like to continue with when imaging an early virtual screening campaign setting. The question was designed so the scientists do not spend much time answering it, relying only on their feelings or \enquote{chemical preference}.

@@

\paragraph{Participants}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree. For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.

\paragraph{Comparison with models}
For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans. The performance metrics reported for models in the main text are computed on the same questions the humans answered. Metrics for the entire corpus are reported in the appendix.

\subsection{Data annotation}\label{sec:meth-topic}
In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question. We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g. the H/P statements of chemicals). Reasoning is required if answering a question requires multiple reasoning steps.
Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.

\section*{Data and code availability}
The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\showyourwork} framework.\autocite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}.
To facilitate reproduction, some intermediate analysis results are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
A.A.\ gratefully acknowledges financial support for this (publication/project/research) by the Fulbright U.S. Student Program, which is sponsored by the U.S. Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.

>>>>

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely accepted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics and skills, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of thousands of questions and answers that we manually or semi-automatically compiled from various sources (see \Cref{sec:curation}).
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions). In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations). We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}

\paragraph{\chembenchmini}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\chembenchmini, encompassing \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human participants answered all the questions in this subset.

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on \chembenchmini.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

Notably, we find that models are still limited in their ability to answer knowledge-intensive questions (\Cref{tab:performance_table_human_subset}), that is they did not memorize the relevant facts. Our results indicate that this is not a limitation that could be overcome by simple application of \gls{rag} systems such as \PaperQATwo. This is likely because the required knowledge cannot easily be accessed via papers (which is the only external knowledge \PaperQATwo has access to) but rather by lookup in specialized databases (e.g., PubChem, Gestis), which also the humans in our study used to answer such questions (\Cref{fig:tool_use}).
This indicates that there is still room for improvement of chemical \gls{llm} by training them on more specialized data sources or by integrating them with specialized databases.

In addition, our analysis shows that the performance of models is correlated with their size (see \Cref{fig:model_size_plot}). This is in line with observations in other domains but also indicates that chemical \gls{llm} could, to some extent, be further improved by scaling them up.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics on \chembenchmini.} The radar plot shows the performance of the models and humans on the different topics of \chembenchmini. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on \chembenchmini. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult even for the best models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_o1.txt} percent correct answers for \oone).
Importantly, while the human experts are given a drawing of the compounds, the models are only shown the \gls{smiles} string of a a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule).

 These findings also shine an interesting light on the value of textbook-inspired questions.
 A subset of the questions in \chembench are based on textbooks targeted at undergraduate students.
 On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
 For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt-4.txt}\% correct answers for \GPTFour, \variable{output/subset_scores/is_gfk_Claude-3.5_(Sonnet).txt}\% for \ClaudeThreeFiveSonnet, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
 While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning or are further away from the training corpus.\autocite{mccoy2023embersautoregressionunderstandinglarge} The findings also underline that such exams might have been a good surrogate for general performance of skills for humans, but their applicability in the face of systems that can consume vast amounts of data is up for debate.

@@

% Estimate of difficulty
\paragraph{Confidence estimates} One might wonder whether the models can estimate if they can answer a question correctly.
If they could do so, incorrect answers would be less problematic.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about the \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

These questions assess the ability to establish a \enquote{preference}, such as favoring a specific molecule. Chemical preference is of major importance in drug discovery projects, where the optimization process to reach the desired molecular properties is a process that takes several years within a chemist's career.
Our data corpus is adapted from the published dataset by \textcite{Choung_2023}, which consists of more than 5000 question-answer pairs about chemical intuition. To build the dataset, they presented 35 medicinal chemists with two different molecules, asking them what molecule they would like to continue with when imaging an early virtual screening campaign setting. The question was designed so the scientists do not spend much time answering it, relying only on their feelings or \enquote{chemical preference}.

@@

\paragraph{Participants}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} has a bachelor's degree. For the analysis, we excluded participants with less than two years of experience in chemistry after their first university-level course in chemistry.

\paragraph{Comparison with models}
For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans. The performance metrics reported for models in the main text are computed on the same questions the humans answered. Metrics for the entire corpus are reported in the appendix (\Cref{sec:model_performance_app}).

\subsection{Data annotation}\label{sec:meth-topic}
In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question. We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g. the H/P statements of chemicals). Reasoning is required if answering a question requires multiple reasoning steps.
Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.

\section*{Data and code availability}
The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\showyourwork} framework.\autocite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}.
To facilitate reproduction, some intermediate analysis results are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S. Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
